\section{Matching games and the proposed method}\label{sec:method}

Both of the initialisation methods described in Section~\ref{sec:inits} have a
greedy component. In the case of Cao's method, it would be entirely greedy if it
did not also attempt to force separation between the set of initial modes. In
the case of Huang's, however, the greediness only comes in at the end of the
method, after \(\textsc{SamplePotentialModes}\). In any practical implementation
of this method, the order in which a set of potential modes is iterated over
will alter the initial set of modes returned; this is down to each allocation
being locally optimal, i.e.\ greedy. The initialisation method proposed in this
work aims to extend Huang's method to be order-invariant in the final allocation
\-- thereby eliminating its greedy component. This is done by constructing and
solving a matching game between the set of potential modes and some subset of
the data.

In general, matching games are defined by two sets of players in which each
player ranks at least some of the players in the other set. The objective then
is to find a mapping between the two sets of players such that no pair of
players is (rationally) unhappy with their mapping.

One of the simplest matching games models the Stable Marriage Problem
(SM)~\cite{Gale1962}. In this game the sets of players must be of equal size and
rank each other strictly (i.e.\ no ties allowed) and entirely. As the motivation
of the proposed method is to avoid greediness, this kind of game was not
considered as it effectively reduces down to Huang's method. That is, each
potential mode gets to pick the data point it is closest to but that hasn't
already been picked. When the game is solved, each potential mode is assigned to
its choice.

The primary motivation for this work is to move away from the greedy approaches
defined above by incorporating some techniques from game theory, namely:
matching games. The purpose of solving a matching game is to link the elements
of two sets in a `fair' way so that no element could feasibly be better off. By
considering the virtual modes found during Huang's method with some suitable
subset of our dataset as a matching game to be solved, we hope to find a
closer-to-optimal set of initial modes for the \(k\)-modes algorithm.

\begin{definition}\label{def:matching_game}
    Consider two sets \(S, R\) each of size \(N\), and let us call them 
    `suitors' and `reviewers'. Each element of \(S\) and \(R\) has associated 
    with it a preference list of the other set's elements. These preference 
    lists are ranked in descending order. We consider the preference lists as 
    functions which produce tuples, \(f\) and \(g\) respectively:
	\[
	    f : S \to R^n, \ g : R \to S^n
	\]
	This construction of sets and preference lists is called a 
    \emph{matching game} of size \(N\) and we denote the game itself by 
    \((S,R)\).
	
    A matching, \(M\), is defined to be any bijection between \(S\) and \(R\). 
    If \(s \in S\) and \(r \in R\) are matched by \(M\), then we write
    \(M(s)~=~r\). A matching \(M\) is considered to be either stable or unstable
    based on the preference lists of its suitors and reviewers, and the presence
    of blocking pairs.
\end{definition}

\begin{definition}\label{def:blocking_pair}
    Let \((S, R)\) be a matching game of size \(N\) with some matching \(M\). A 
    pair \((s, r) \in S \times R\) is said to \emph{block} \(M\) if \(M(s) \neq
    r\) but \(s\) prefers \(r\) to \(M(s) = r'\) and \(r\) prefers \(s\) to
    \(M^{-1}(r) = s'\). That is, \(r\) appears before \(r'\) in \(f(s)\) and
    \(s\) appears before \(s'\) in \(g(r)\).
\end{definition}

\begin{definition}\label{def:stable_matching}
    Let \((S, R)\) be a matching game of size \(N\) with some matching \(M\). 
    Then we say \(M\) is a \emph{stable matching} if there are no blocking 
    pairs, and \emph{unstable} otherwise.
\end{definition}

%\input{tex/examples/matching.tex}

\subsection{The Gale-Shapley algorithm}\label{subsec:galeshapley}

The Gale-Shapley algorithm is known to find a unique stable matching for any 
matching game of size \(N\). This matching is also considered to be 
suitor-optimal. That is, each suitor is matched with the best possible reviewer
that ensures a stable matching, but is in fact the worst possible matching for 
the reviewers. \textcolor{red}{[cite or have these theorems stated/proven]} 

As was discussed at the start of Section~\ref{sec:matching}, the outline of
the method proposed in this paper is to extend Huang's method by considering the
virtual modes with some subset of the data as a matching game to be solved. It
should be noted, however, that in this method we may not have equally sized 
sets of suitors and reviewers. As a result of this, the Gale-Shapley algorithm 
becomes inapplicable as the matching produced \(M\) would not be a bijection of 
our suitors and reviewers.

\subsection{The capacitated Gale-Shapley algorithm for the hospital-resident 
	problem}\label{subsec:capacitated_galeshapley}

The situation where a large set of suitors are to be matched with a number
reviewers is not limited to abstraction. A practical example of this problem is
how to best assign a cohort of medical students to a group of hospitals. Here, 
we have all of the requisite components of a matching game:

\begin{itemize}
	\item A set of reviewers (the hospitals) and a set of suitors (the potential
        residents) 
	\item A ranking of the students/residents by the hospitals, and vice versa
\end{itemize}

The only obstacle which stops us from using the Gale-Shapley algorithm is the 
disparity in the sizes of our sets. In reality, hospitals need not always take 
at most one resident on from a cohort of medical students. So each hospital has
a capacity associated with it and we can consider our matching game to be
`capacitated'. By this we mean that each reviewer (hospital) may be matched with
any number of suitors (students) up to their capacity, making our matching
\(M:~S~\to~R\) surjective.

Research surrounding the hospital-resident assignment problem is well-documented 
\textcolor{red}{[cite]} and an extension of the Gale-Shapley algorithm was
developed to solve it, awarding the authors the 2012 Nobel Prize in Economic
Sciences. This algorithm is currently used by the National Resident Matching
Program (\url{http://www.nrmp.org}) to assign medical students to hospitals in
the United States of America.

As before, we consider a set of suitors and reviewers denoted by \(S\) and 
\(R\). These sets are no longer (necessarily) the same size. We also have our 
preference lists \(f, g\), and a set \(C = \{c_1, \ldots, c_{|R|}\}\) of 
reviewer capacities. Finally, let \(S_u \subset S\) denote the set of suitors 
that are currently unmatched. The capacitated Gale-Shapley algorithm is given 
below.

\input{tex/algorithms/hospital_resident.tex}
